---
title: "Data Visulizaton Final Project"
author: Junling Shao, Yuhan Ma
editor: visual
format: 
  html: 
    embed-resources: true
---

Conflicts often have profound social, economic, and political consequences, and understanding their distribution and trends is crucial for effective decision-making and resource allocation. The overview provides an **overview of global conflict activity and its relationship to unemployment trends**, helping us answer key questions like:

-   Which regions have experienced consistently high levels of conflict over time?

-   How has the distribution of conflict evolved from 1990 to 2023?

-   Are there emerging regions that require more attention due to rising conflict trends?

By exploring these patterns, we aim to identify focus areas for deeper analysis and intervention.

## Part 1: Stacked Area Chart

The **stacked area chart** is the first step in our analysis, offering a bird's-eye view of **conflict distribution across regions over time**. It visualizes the total number of conflicts globally, broken down into regional contributions from 1990 to 2023. The chart serves as a foundation to contextualize the scale and dynamics of conflict activity.

```{r}
#| echo: false
#| warning: false
#| message: false
setwd('/Users/machuichui/Desktop/2024fall/DV/final project')

library(tidyverse)
library(tidyr)
library(ggplot2)
library(dbplyr)
library(broom)
library(lubridate)
library(countrycode)
library(sf)
library(openxlsx)
library(plotly)
library(RColorBrewer)
library(viridis)
library(rnaturalearth)

unemp_sect <- read.csv("ACLED-sector-employment-annual-2000-2023.csv") # Sector unemp
unemp_dem <- read.csv("ACLED-unemployment-annual-1997-2023.csv") # demographic unemp
acled <- read.csv("ACLED-event-level-1997-2024.csv") # ACLED 
region_code = read.xlsx('sub_region_code.xlsx')
```

```{r}
#| echo: false
#| warning: false
unemp_dem <- unemp_dem %>%
  left_join(region_code %>% select(country_code, sub_region_name), by = "country_code") %>%
  rename(sub_region = sub_region_name) # Rename column if needed

# Add `sub_region_name` to unemp_sect
unemp_sect <- unemp_sect %>%
  left_join(region_code %>% select(country_code, sub_region_name), by = "country_code") %>%
  rename(sub_region = sub_region_name) # Rename column if needed
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Step 1: Exclude rows where `acled_num_incidents_in_year` is NA
unemp_dem_clean <- unemp_dem %>%
  filter(!is.na(acled_num_incidents_in_year))

# Step 2: Generate a dynamic color palette
# Use RdYlBu for professional-looking colors
color_palette <- hcl.colors(n = length(unique(unemp_dem_clean$sub_region)), palette = "RdYlBu")

# Step 3: Aggregate and calculate percentages
sub_region_data <- unemp_dem_clean %>%
  group_by(year, sub_region) %>%
  summarise(
    sub_region_total = sum(acled_num_incidents_in_year, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(year) %>%
  mutate(
    global_total = sum(sub_region_total, na.rm = TRUE),
    percentage_contribution = (sub_region_total / global_total) * 100 # Keep as percentage
  ) %>%
  ungroup()

# Step 4: Create the 100% stacked bar chart
interactive_plot <- ggplot(sub_region_data, aes(x = year, y = percentage_contribution, fill = sub_region)) +
  geom_bar(stat = "identity", position = "fill", aes(
    text = paste(
      "Year:", year, "<br>",
      "Sub-Region:", sub_region, "<br>",
      "Incidents:", sub_region_total, "<br>",
      "Percentage", sprintf("%.1f%%", percentage_contribution)
    )
  )) +
  scale_fill_manual(values = color_palette) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Global Conflict Activity by Sub-Region",
    x = "Year",
    y = "Percentage",
    fill = "Sub-Region"
  ) +
  theme_minimal()

# Step 5: Convert ggplot to plotly and apply highlight settings
interactive_plot <- ggplotly(interactive_plot, tooltip = "text") %>%
  highlight(
    on = "plotly_click",
    persistent = TRUE    
  )

# Display the interactive plot
interactive_plot
```

## Part 2: Linked Chart (Heatmap and Line Chart)

[Click here to open the Shiny App](https://js5241.shinyapps.io/viz_final/)

## Part3. Relationship between Conflicts and unemployment Rate By Demographic in Selected Regions

From the previous analysis, we have decided to focus on **Sub-Saharan Africa** and **Southern Asia**. In this section, we will delve deeper into how conflicts impact unemployment across different gender and youth groups in these two areas.

```{r}
#| echo: false
#| warning: false
# create the dataset that only have the unemployment for female and male by sub region.
prepared_data_dem_gender <- unemp_dem %>%
  filter(
    unemployment_metric %in% c(
      "Unemployment, female (% of female labor force) (modeled ILO estimate)",
      "Unemployment, male (% of male labor force) (modeled ILO estimate)"
    )
  ) %>%
  group_by(year, sub_region, unemployment_metric) %>%  
  summarise(
    unemployment_value = mean(unemployment_value, na.rm = TRUE),  
    num_incidents_per100k_people = mean(num_incidents_per100k_people, na.rm = TRUE),  
    .groups = "drop"  
  ) %>%
  pivot_wider(
    names_from = unemployment_metric,
    values_from = unemployment_value
  ) %>%
  rename(
    unemployment_rate_female = `Unemployment, female (% of female labor force) (modeled ILO estimate)`,
    unemployment_rate_male = `Unemployment, male (% of male labor force) (modeled ILO estimate)`
  ) %>%
  select(
    year,
    sub_region,
    unemployment_rate_female,
    unemployment_rate_male,
    num_incidents_per100k_people
  )


```

### Sub-Saharan Africa

This dataset for this part was prepared by first filtering the raw data to include only metrics related to unemployment rates for females and males as a percentage of their respective labor forces (modeled ILO estimates). After filtering, the data was grouped by year, sub-region, and unemployment metric. Within each group, the unemployment values and the number of incidents per 100,000 people were aggregated by calculating their mean, excluding missing values. The dataset was then pivoted to create separate columns for female and male unemployment rates, which were appropriately renamed for clarity. Finally, the dataset was organized to retain only the year, sub-region, female unemployment rate, male unemployment rate, and the number of incidents per 100,000 people. Same as the dataset for Southern Asia.

The figure analyzes the relationship between conflicts and unemployment across gender groups in Sub-Saharan Africa. Key observations include：

Conflict Intensity:

-   The green curve representing "Incidents per 100,000 People" reveals a significant escalation in conflict intensity from the early 2000s

-   Conflict peaks occurred around 2015-2020

Unemployment Dynamics:

-   Shaded areas depict unemployment rates for males and females over time

-   The red dashed line illustrates the gender-based unemployment disparity. Females consistently experienced higher unemployment rates compared to males

-   The unemployment rate gap fluctuates without a clear systematic pattern

```{r}
#| echo: false
#| warning: false

# Filter the data for Sub-Saharan Africa and calculate Female - Male difference
filtered_data_dem_gender_subshr <- prepared_data_dem_gender %>%
  filter(sub_region == "Sub-Saharan Africa") %>%
  mutate(difference = unemployment_rate_female - unemployment_rate_male)  


plot_ly() %>%
  # Add Female Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_gender_subshr,
    x = ~year,
    y = ~unemployment_rate_female,
    type = 'scatter',
    mode = 'lines',  # Use 'lines' mode
    name = 'Unemployment Rate (Female)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',  
    line = list(shape = 'spline')
  ) %>%
  # Add Male Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_gender_subshr,
    x = ~year,
    y = ~unemployment_rate_male,
    type = 'scatter',
    mode = 'lines',
    name = 'Unemployment Rate (Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',  
    line = list(shape = 'spline')
  ) %>%
  # Add Incidents per 100k People
  add_trace(
    data = filtered_data_dem_gender_subshr,
    x = ~year,
    y = ~num_incidents_per100k_people,
    type = 'scatter',
    mode = 'lines',
    fill = 'tozeroy',
    name = 'Incidents per 100k People',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',   
    line = list(shape = 'spline'),
    yaxis = 'y2'
  ) %>%
  # Add Difference Line (Female - Male)
  add_trace(
    data = filtered_data_dem_gender_subshr,
    x = ~year,
    y = ~difference,
    type = 'scatter',
    mode = 'lines',
    name = 'Difference (Female - Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',  
    line = list(shape = 'spline', dash = 'dot'),
    yaxis = 'y'
  ) %>%
  layout(
    title = list(
      text = "Relationship between Conflicts Numbers and Unemployment Rate \nin Sub-Saharan Africa by Gender",
      font = list(size = 15)
    ),
    xaxis = list(
      title = "Year",
      tickangle = 45,
      rangeslider = list(visible = TRUE),
      rangeselector = list(
        buttons = list(
          list(count = 5, label = "Last 5 Years", step = "year", stepmode = "backward"),
          list(count = 10, label = "Last 10 Years", step = "year", stepmode = "backward"),
          list(step = "all", label = "All Time")
        )
      )
    ),
    yaxis = list(
      title = "Unemployment Rate (%)",
      side = "left",
      range = c(0, 20)
    ),
    yaxis2 = list(
      title = "Incidents per 100k People",
      overlaying = "y",
      side = "right"
    ),
    legend = list(
      x = 1.15,
      y = 1,
      xanchor = "left"
    )
  )


```

```{r}
#| echo: false
#| warning: false

prepared_data_dem_youth <- unemp_dem %>%
  filter(
    unemployment_metric %in% c(
      "Unemployment, youth female (% of female labor force ages 15-24) (modeled ILO estimate)",
      "Unemployment, youth male (% of male labor force ages 15-24) (modeled ILO estimate)"
    )
  ) %>%
  group_by(year, sub_region, unemployment_metric) %>%  
  summarise(
    unemployment_value = mean(unemployment_value, na.rm = TRUE),  
    num_incidents_per100k_people = mean(num_incidents_per100k_people, na.rm = TRUE),  
    .groups = "drop"  
  ) %>%
  pivot_wider(
    names_from = unemployment_metric,
    values_from = unemployment_value
  ) %>%
  rename(
    unemployment_rate_youth_female = `Unemployment, youth female (% of female labor force ages 15-24) (modeled ILO estimate)`,
    unemployment_rate_youth_male = `Unemployment, youth male (% of male labor force ages 15-24) (modeled ILO estimate)`
  ) %>%
  select(
    year,
    sub_region,
    unemployment_rate_youth_female,
    unemployment_rate_youth_male,
    num_incidents_per100k_people
  )


```

The chart examines youth employment dynamics in Sub-Saharan Africa during periods of conflict. Compared to the total population, the unemployment gap between young women and men appears even more significant. This underscores the heightened economic vulnerability of young women in regions experiencing social and political instability.

```{r}
#| echo: false
#| warning: false

# Filter the data for Sub-Saharan Africa and calculate Female - Male difference
filtered_data_dem_youth_subshr <- prepared_data_dem_youth %>%
  filter(sub_region == "Sub-Saharan Africa") %>%
  mutate(difference = unemployment_rate_youth_female - unemployment_rate_youth_male)  


plot_ly() %>%
  # Add Female Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_youth_subshr ,
    x = ~year,
    y = ~unemployment_rate_youth_female,
    type = 'scatter',
    mode = 'lines',  # Use 'lines' mode
    name = 'Unemployment Rate (Youth Female)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline')  # Smooth curve
  ) %>%
  # Add Youth Male Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_youth_subshr,
    x = ~year,
    y = ~unemployment_rate_youth_male,
    type = 'scatter',
    mode = 'lines',  
    name = 'Unemployment Rate (Youth Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline') 
  ) %>%
  # add Youth Female Unemployment Rate As a Line
  add_trace(
    data = filtered_data_dem_youth_subshr ,
    x = ~year,
    y = ~num_incidents_per100k_people,
    type = 'scatter',
    mode = 'lines',  
    fill = 'tozeroy',  
    name = 'Incidents per 100k People',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline'),  
    yaxis = 'y2'  
  ) %>%
  # Add Difference Line (Female - Male)
  add_trace(
    data = filtered_data_dem_youth_subshr,
    x = ~year,
    y = ~difference,
    type = 'scatter',
    mode = 'lines',  
    name = 'Difference (Youth Female - Youth Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline', dash = 'dot'),  
    yaxis = 'y' 
  ) %>%
  layout(
    title = list(
      text = "Relationship between Conflicts Numbers and Unemployment Rate \nin Sub-Saharan Africa by Youth Gender",
      font = list(size = 15)  
    ),
    xaxis = list(
      title = "Year",
      tickangle = 45,  
      rangeslider = list(visible = TRUE),  
      rangeselector = list(  
        buttons = list(
          list(count = 5, label = "Last 5 Years", step = "year", stepmode = "backward"),
          list(count = 10, label = "Last 10 Years", step = "year", stepmode = "backward"),
          list(step = "all", label = "All Time")
        )
      )
    ),
    yaxis = list(
      title = "Unemployment Rate (%)",
      side = "left"
    ),
    yaxis2 = list(
      title = "Incidents per 100k People",
      overlaying = "y",  
      side = "right"
    ),
    legend = list(
      x = 1.15,  
      y = 1,  
      xanchor = "left"
    )
  )
  
```

### Southern Asia

The chart provides insights into the impact of conflicts on employment in Southern Asia. It shows that

-   conflict intensity, as measured by incidents per 100,000 people, has increased significantly since the early 2000s, peaking around 2015-2020.

-   The data also reveals persistent gender disparities in unemployment rates, with females consistently experiencing higher unemployment compared to males. Importantly, the analysis notes that the gender gap in unemployment is even more pronounced after the peak of conflicts

```{r}
#| echo: false
#| warning: false

# Filter the data for Southern Asia and calculate Female - Male difference
filtered_data_dem_gender_sa <- prepared_data_dem_gender %>%
  filter(sub_region == "Southern Asia") %>%
  mutate(difference = unemployment_rate_female - unemployment_rate_male)  

plot_ly() %>%
  # Add Female Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_gender_sa,
    x = ~year,
    y = ~unemployment_rate_female,
    type = 'scatter',
    mode = 'lines',  # Use 'lines' mode
    name = 'Unemployment Rate (Female)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline')  
  ) %>%
  # Add Male Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_gender_sa,
    x = ~year,
    y = ~unemployment_rate_male,
    type = 'scatter',
    mode = 'lines',  
    name = 'Unemployment Rate (Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline')  
  ) %>%
  
  add_trace(
    data = filtered_data_dem_gender_sa,
    x = ~year,
    y = ~num_incidents_per100k_people,
    type = 'scatter',
    mode = 'lines',  
    fill = 'tozeroy', 
    name = 'Incidents per 100k People',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline'), 
    yaxis = 'y2'  
  ) %>%
  # Add Difference Line (Female - Male)
  add_trace(
    data = filtered_data_dem_gender_sa,
    x = ~year,
    y = ~difference,
    type = 'scatter',
    mode = 'lines', 
    name = 'Difference (Female - Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline', dash = 'dot'),  
    yaxis = 'y'  
  ) %>%
  layout(
    title = list(
      text = "Relationship between Conflicts Numbers and Unemployment Rate in Southern Asia by Gender",
      font = list(size = 14) 
    ),
    xaxis = list(
      title = "Year",
      tickangle = 45,  
      rangeslider = list(visible = TRUE),  
      rangeselector = list(
        buttons = list(
          list(count = 5, label = "Last 5 Years", step = "year", stepmode = "backward"),
          list(count = 10, label = "Last 10 Years", step = "year", stepmode = "backward"),
          list(step = "all", label = "All Time")
        )
      )
    ),
    yaxis = list(
      title = "Unemployment Rate (%)",
      side = "left",
      range = c(0, 25) 
    ),
    yaxis2 = list(
      title = "Incidents per 100k People",
      overlaying = "y",  
      side = "right"
    ),
    legend = list(
      x = 1.15,  
      y = 1,  
      xanchor = "left"
    )
  )
```

The impact on youth group in Southern Asia has

-   The increasing trend of unemployment in youth group is more obvious than that in total population.

-   The female youth unemployment rate is consistently higher than the male rate.

-   The unemployment gap between youth male and female seems to slightly narrow down compared with total group.

```{r}
#| warning: false
#| echo: false
# Filter the data for Southern Asia and calculate Female - Male difference
filtered_data_dem_youth_sa <- prepared_data_dem_youth %>%
  filter(sub_region == "Southern Asia") %>%
  mutate(difference = unemployment_rate_youth_female - unemployment_rate_youth_male) 

plot_ly() %>%
  # Add Female Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_youth_sa,
    x = ~year,
    y = ~unemployment_rate_youth_female,
    type = 'scatter',
    mode = 'lines',  
    name = 'Unemployment Rate (Youth Female)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline')  
  ) %>%
  # Add Youth Male Unemployment Rate as a line
  add_trace(
    data = filtered_data_dem_youth_sa,
    x = ~year,
    y = ~unemployment_rate_youth_male,
    type = 'scatter',
    mode = 'lines',  
    name = 'Unemployment Rate (Youth Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline')  
  ) %>%
  # add Youth Female Unemployment Rate As a Line
  add_trace(
    data = filtered_data_dem_youth_sa,
    x = ~year,
    y = ~num_incidents_per100k_people,
    type = 'scatter',
    mode = 'lines',  
    fill = 'tozeroy',  
    name = 'Incidents per 100k People',
    hoverinfo = 'x+y+name',
    line = list(shape = 'spline'),  
    yaxis = 'y2'  
  ) %>%
  # Add Difference Line (Female - Male)
  add_trace(
    data = filtered_data_dem_youth_sa,
    x = ~year,
    y = ~difference,
    type = 'scatter',
    mode = 'lines',  
    name = 'Difference (Youth Female - Youth Male)',
    hovertemplate = 'Year: %{x}<br>Rate: %{y:.2f}<extra></extra>',
    line = list(shape = 'spline', dash = 'dot'),  
    yaxis = 'y' 
  ) %>%
  layout(
    title = list(
      text = "Relationship between Conflicts Numbers and Unemployment Rate in Southern Asia by Youth Gender",
      font = list(size = 15)  
    ),
    xaxis = list(
      title = "Year",
      tickangle = 45, 
      rangeslider = list(visible = TRUE), 
      rangeselector = list(  
        buttons = list(
          list(count = 5, label = "Last 5 Years", step = "year", stepmode = "backward"),
          list(count = 10, label = "Last 10 Years", step = "year", stepmode = "backward"),
          list(step = "all", label = "All Time")
        )
      )
    ),
    yaxis = list(
      title = "Unemployment Rate (%)",
      side = "left"
    ),
    yaxis2 = list(
      title = "Incidents per 100k People",
      overlaying = "y", 
      side = "right"
    ),
    legend = list(
      x = 1.15,  
      y = 1,  
      xanchor = "left"
    )
  )
```

An interesting difference in their impacts lies in the timing of the peaks in unemployment relative to the peak of conflicts. In Southern Asia, unemployment rates peak after the peak in conflicts, suggesting a delayed effect. Conversely, Sub-Saharan Africa, unemployment peaks immediately after the peak in conflicts, indicating a more immediate impact.

## Part 4: Relationship between Conflicts and Employment Increase Rate By Sector in Selected Regions

In this section, we continue exploring the impact of conflicts on employment in Sub-Saharan Africa and Southern Asia. The dataset provides the percentage of the total population employed in each sector, using whole population in nation as denominator, which may not reflect sector-specific employment demand. To address this, we calculated the **employment increase rate** as a more reasonable index. 


We calculates the employment increase rate by grouping the data by country_name and industry_sector, and then computing the year-over-year percentage change in the proportion of the population employed in each sector,simply:
$$
\text{Employment Increase Rate} = \frac{\text{Current Year's Percentage} - \text{Previous Year's Percentage}}{\text{Previous Year's Percentage}}
$$
Afterward, we filters the data for the "Sub-Saharan Africa" region and aggregates it further by year, sub_region, and industry_sector. For each group, it calculates the average employment increase rate and the average number of incidents per 100,000 people.


```{r}
#| echo: false
#| warning: false
unemp_sect <- unemp_sect %>%
  arrange(country_name, industry_sector, year) %>%  
  group_by(country_name, industry_sector) %>%
  mutate(
    employment_increase_rate = ifelse(
      is.na(lag(pct_total_pop_employed_in_sector)), 
      NA,  
      (pct_total_pop_employed_in_sector - lag(pct_total_pop_employed_in_sector)) / lag(pct_total_pop_employed_in_sector)
    )
  ) %>%
  ungroup()  


filtered_data_sec_subshr <- unemp_sect %>%
  filter(sub_region == "Sub-Saharan Africa") %>%  
  group_by(year, sub_region, industry_sector) %>%
  summarise(
    employment_increase_rate = mean(employment_increase_rate, na.rm = TRUE), 
    num_incidents_per100k_people = mean(num_incidents_per100k_people, na.rm = TRUE),  
    .groups = "drop"  
  )



```

```{r}
#| warning: false
#| echo: false
sector_correlation_subshr <- filtered_data_sec_subshr %>%
  filter(
    !is.na(employment_increase_rate) & 
    !is.na(num_incidents_per100k_people)
  ) %>%  # Drop rows with NA in the relevant columns
  group_by(industry_sector) %>%
  filter(n() >= 2) %>% 
  summarise(
    correlation = cor(
      employment_increase_rate, 
      num_incidents_per100k_people, 
      use = "complete.obs"  # Ensures only complete pairs are used
    ),
    .groups = "drop"
  )





```



### Sub-Saharan Africa


This bar chart highlights the correlation between incidents and the employment increase rate across various industry sectors in Sub-Saharan Africa. Positive correlations are observed in sectors like "Shipping, ports, fisheries, inland waterways," indicating that conflict-related incidents may coincide with increases in employment in these areas. Conversely, negative correlations are found in sectors such as "Public service" and "Forestry, wood, pulp, and paper," suggesting that these industries might experience reduced employment growth in the context of conflicts.

```{r}
#| warning: false
#| echo: false
ggplot(sector_correlation_subshr, aes(x = reorder(industry_sector, correlation), y = correlation, fill = correlation)) + 
  geom_bar(stat = "identity") + 
  coord_flip() +  
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Correlation" 
  ) + 
  labs(
    title = "Correlation Between Incidents and Employment \nIncrease Rate by Sector in Sub-Saharan Africa",
    x = "Industry Sector",
    y = "Correlation"
  ) + 
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold", margin = margin(b = 15)),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10),  
    axis.text.y = element_text(size = 10),  
    legend.position = "right",  
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)  
  )

```

With the animated buble chart, we can observe that from 2011-2015, the employment increase rate in basic metal production jumped out.However, with the increasing conflicts after 2016, the oil,gas industry were hiring more and more people.

```{r}
#| echo: false
#| warning: false
filtered_data_sec_subshr <- filtered_data_sec_subshr %>%
  filter(industry_sector != "Not elsewhere classified")

# Generate a color palette using viridis
unique_sectors <- unique(filtered_data_sec_subshr$industry_sector)
num_sectors <- length(unique_sectors)
sector_colors <- viridis(num_sectors, option = "D")

fig <- plot_ly(
  data = filtered_data_sec_subshr,
  x = ~num_incidents_per100k_people,
  y = ~employment_increase_rate,
  size = ~employment_increase_rate,
  color = ~industry_sector,
  colors = sector_colors,
  frame = ~year,  # Set the frame for animation
  text = ~paste(
    "Sector:", industry_sector, "<br>",
    "Year:", year, "<br>",
    "Incidents per 100k People:", round(num_incidents_per100k_people, 2), "<br>",
    "Employment Increase Rate:", round(employment_increase_rate, 2)
  ),
  hoverinfo = "text",
  type = 'scatter',
  mode = 'markers',
  marker = list(sizemode = 'diameter')
) %>%
  layout(
    title = "Employment Increase Rate by Sector Over Years in Sub-Saharan Africa",
    xaxis = list(title = "Incidents per 100k People"),
    yaxis = list(title = "Employment Increase Rate"),
    showlegend = TRUE,
    updatemenus = list(
      list(
        type = "buttons",
        showactive = TRUE,
        x = -0.2, 
        y = 1, 
        buttons = list(
          list(  # Custom pause button
            label = "Pause",
            method = "animate",
            args = list(NULL, list(frame = list(duration = 0, redraw = FALSE), mode = "immediate"))
          )
        )
      )
    )
  )


fig
```



```{r}
#| echo: false
#| warning: false
filtered_data_sec_sa <- unemp_sect %>%
  filter(sub_region == "Southern Asia") %>%   
  group_by(year, sub_region, industry_sector) %>%
  summarise(
    employment_increase_rate = mean(employment_increase_rate, na.rm = TRUE), 
    num_incidents_per100k_people = mean(num_incidents_per100k_people, na.rm = TRUE),  
    .groups = "drop"  
  )


```

```{r}
#| echo: false
#| warning: false
sector_correlation_sa <- filtered_data_sec_sa %>%
  group_by(industry_sector) %>%
  summarise(
    correlation = cor(
      employment_increase_rate, 
      num_incidents_per100k_people, 
      use = "complete.obs"  
    ),
    .groups = "drop"
  )


```

### Southern Asia

The employment increase rate and data aggregation methods are the same as those used for the Sub-Saharan Africa region.

This chart presents the correlation between conflict incidents and the employment increase rate across industry sectors in Southern Asia. Positive correlations in sectors like "Oil and gas production, oil refining" and "Health services" suggest that these industries experience employment growth during periods of higher conflict. Conversely, negative correlations, notably in "Transport equipment manufacturing" and "Forestry, wood, pulp, and paper," indicate that employment in these sectors tends to decline as conflict incidents increase. And compared with Sub-Saharan Africa, conflicts seemed to bring more negative impact on employment in Southern Asia.

```{r}
#| echo: false
#| warning: false
ggplot(sector_correlation_sa, aes(x = reorder(industry_sector, correlation), y = correlation, fill = correlation)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Correlation Between Incidentsand Employment \nIncrease Rate by Sector in Southern Asia",
    x = "Industry Sector",
    y = "Correlation"
  ) +
  theme_minimal() +
   theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold", margin = margin(b = 15)),  # Center title and add spacing below
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10),  
    axis.text.y = element_text(size = 10),  
    legend.position = "right",  
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20) 
  )
```

From the bubble chart, we can observe an employment increase in the construction industry before 2015. However, after 2015, with the rise in conflicts, few industries demonstrated strong performance in employment increase rates.

```{r}
#| echo: false
#| warning: false
filtered_data_sec_sa <- filtered_data_sec_sa %>%
  filter(industry_sector != "Not elsewhere classified")

# Generate a color palette using viridis
unique_sectors <- unique(filtered_data_sec_sa$industry_sector)
num_sectors <- length(unique_sectors)
sector_colors <- viridis(num_sectors, option = "D")

fig <- plot_ly(
  data = filtered_data_sec_sa,
  x = ~num_incidents_per100k_people,
  y = ~employment_increase_rate,
  size = ~employment_increase_rate,
  color = ~industry_sector,
  colors = sector_colors,
  frame = ~year,  # Set the frame for animation
  text = ~paste(
    "Sector:", industry_sector, "<br>",
    "Year:", year, "<br>",
    "Incidents per 100k People:", round(num_incidents_per100k_people, 2), "<br>",
    "Employment Increase Rate:", round(employment_increase_rate, 2)
  ),
  hoverinfo = "text",
  type = 'scatter',
  mode = 'markers',
  marker = list(sizemode = 'diameter')
) %>%
  layout(
    title = "Employment Increase Rate by Sector Over Years in Southern Asia",
    xaxis = list(title = "Incidents per 100k People"),
    yaxis = list(title = "Employment Increase Rate"),
    showlegend = TRUE,
    updatemenus = list(
      list(
        type = "buttons",
        showactive = TRUE,
        x = -0.2,  
        y = 1,  
        buttons = list(
          list(  
            label = "Pause",
            method = "animate",
            args = list(NULL, list(frame = list(duration = 0, redraw = FALSE), mode = "immediate"))
          )
        )
      )
    )
  )

# Display the figure
fig
```
